---
title: "covid19"
authors: Diane Kim, Arielle Hutchinson, Priya Parkash
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include = FALSE}
library(tidyverse)
library("dplyr")
library(stringr)
library(RCurl)
library(betareg)
```

```{r load-data, include = FALSE}
date_of_study = "04-20-2020"

covid_hist = read.csv(text=getURL("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-30-2020.csv"))
covid_us_hist = subset(covid_hist, Country_Region == "US" & is.na(FIPS)==F)

covid = read.csv(text=getURL(paste0("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/",date_of_study,".csv")))
covid_us = subset(covid,Country_Region == "US" & is.na(FIPS)!=T)
covid_us = rbind(covid_us,subset(covid_us_hist, (!(FIPS %in% covid_us$FIPS))  & Confirmed == 0 & Deaths == 0 & is.na(FIPS)==F))

county_pm = read.csv(text=getURL("https://raw.githubusercontent.com/wxwx1993/PM_COVID/master/Data/county_pm25.csv"))

pm_ave <- read_csv("county_pm_ave.csv")

census <- read_csv("census_county_interpolated.csv")

census_ave <- read_csv("census_ave.csv")
```

```{r clean-up-data}
covid_us <- covid_us %>%
  mutate(fips = FIPS)

pm_mean = county_pm %>% 
  group_by(fips) %>% 
  summarise(pm25 = mean(pm25))
  
```

```{r calculate-mortality}
covid_us <- covid_us %>%
  group_by(fips) %>%
  mutate(mortality = (1+Deaths)/Confirmed) %>%
  filter(mortality<1 & !is.na(mortality))
```

```{r merge-datasets}
fullset <- inner_join(covid_us, pm_mean, by = "fips")
ggplot(data = fullset, mapping = aes(x = mortality))+
  geom_histogram()

full <- inner_join(covid_us, census_ave, by = "fips")
full <- full %>%
  mutate(poverty = 100*poverty, medianhouseholdincome = medianhouseholdincome/1000, medianhousevalue = medianhousevalue/1000, pct_blk = pct_blk*100, pct_asian = pct_asian*100, pct_native = pct_native*100, hispanic = hispanic*100, pct_white = pct_white*100, education = education*100, pct_owner_occ = pct_owner_occ*100)
```

## Exploratory Data Analysis

## Mortality Rates & Race
```{r eda-mortality}
ggplot(data = full, mapping = aes(x = mortality))+
  geom_histogram(fill = "#15246b" )+
  labs(title = "Distribution of Mortality Rate", x = "Mortality", y = "Frequency") +
  theme_minimal()

ggplot(data = full, mapping = aes(x = hispanic, y = mortality)) +
  geom_point(color = "blue")+
  labs(title = "The Mortality Rates for Hispanics", 
       x = "Percent of Hispanic People", y = "Mortality Rate") +
theme_minimal()

ggplot(data = full, mapping = aes(x = pct_blk, y = mortality)) +
  geom_point(color = "lightblue")+
  labs(title = "The Mortality Rates for Blacks", 
       x = "Percent of Blacks", y = "Mortality Rate") +
  theme_minimal()

ggplot(data = full, mapping = aes(x = pct_native, y = mortality) )+
  geom_point(color = "darkblue")+
  labs(title = "The Mortality Rates for Natives", 
       x = "Percent of Natives", y = "Mortality Rate") +
  theme_minimal()

```

im confused. why do these all look so wierd??? also what should the title of these be?


## Mortality, Home-ownership, & Income

```{r eda-household}
ggplot(data = full, mapping = aes(x = medianhouseholdincome))+
  geom_histogram(fill = "darkblue")+
  labs(title = "Distribution of Median Household Income", 
       x = "Household Income", y = "Frequency") +
  theme_minimal()

ggplot(data = full, mapping = aes(x = medianhousevalue)) +
  geom_histogram(fill = "lightblue") +
  labs(title = "Distribution of Median House Value", 
       x = "Median House Value", y = "Frequency") +
  theme_minimal()

ggplot(data = full, mapping = aes(x = pct_owner_occ, y = mortality)) +
  geom_point(color = "blue")+
  labs(title = "The Relationship between Mortality and Owner-Occupied Homes", 
       x = "Percentage of Owner-Occupied Homes", y = "Mortality") +
  theme_minimal()

ggplot(data = full, mapping = aes(x = medianhousevalue, y = mortality)) +
  geom_point(color = "lightblue")+
  labs(title = "The Relationship between Mortality and Median House Value", 
         x = "Median Household Value", y = "Mortality") +
  theme_minimal()

ggplot(data = full, mapping = aes(x = medianhouseholdincome, y = mortality)) +
  geom_point(color = "darkblue")+
  labs(title = "The Mortality Rates for Blacks", 
        x = "Median Household Income", y = "Mortality") +
  theme_minimal()
```

## Poverty & Race
```{r eda-poverty}
ggplot(data = full, mapping = aes(x = poverty))+
  geom_histogram(fill = "#15246b") +
  labs(title = "Distribution of Poverty Rate", 
       x = "Poverty Rate (in percentages)", y = "Frequency") +
  theme_minimal()

ggplot(data = full, mapping = aes(x = pct_white))+
  geom_histogram(fill = "mediumblue") +
  labs(title = "Distribution of White Population", 
       x = "Percentage of Whites", y = "Frequency") +
  theme_minimal()

ggplot(data = full, mapping = aes(x = pct_blk))+
  geom_histogram(fill="gray") +
  labs(title = "Distribution of Black Population", 
       x = "Percentage of Blacks", y = "Frequency") +
  theme_minimal()

ggplot(data = full, mapping = aes(x = hispanic, y = poverty))+
  geom_point(color = "mediumblue")+
  labs(title = "Relationship between Poverty Rate and Hispanic Population", 
       x = "Percentage of Hispanics", y = "Poverty Rate") +
  theme_minimal()

ggplot(data = full, mapping = aes(x = pct_blk, y = poverty))+
  geom_point(color = "gray")+
  labs(title = "Relationship between Poverty Rate and Black Population", 
       x = "Percentage of Blacks", y = "Poverty Rate") +
  theme_minimal()

ggplot(data = full, mapping = aes(x = pct_native, y = poverty))+
  geom_point(color = "darkblue")+
  labs(title = "Relationship between Poverty Rate and Native Population", 
       x = "Percentage of Natives", y = "Poverty Rate") +
  theme_minimal()
```

There seems to be a positive correlation between the percentage of African Americans and the poverty rate. Additionally, there seems to be positive correlation between the percentage of Hispanics and the poverty rate.

## Mortality & Poverty
```{r mortality-poverty}
ggplot(data = full, mapping = aes(x = poverty, y = mortality))+
  geom_point(color = "darkblue")+
  labs(title = "Relationship between Poverty Rate and Mortality Rate", 
       x = "Poverty Rate", y = "Mortality Rate") +
  theme_minimal()

```


## Regression model

```{r betareg}
mod <- betareg(mortality ~ poverty + medianhouseholdincome + medianhousevalue + pct_asian + pct_white + pct_native + education + popdensity + pct_owner_occ + pct_white*education + poverty*pct_blk +poverty*pct_native, data = full)
tidy(mod) %>%
  kable(format = "markdown", digits = 6)
```
The most significant predictor variables are poverty, medianhouseholdincome, medianhousevalue, pct_native, and pct_owner_occ. 

For every 1 percent increase in the poverty rate, we expect the odds of a confirmed patient dying to increase by a multiplicative factor of exp(0.031076), holding all else constant.
